A Formal Model of Checked C

### 2022-08-01 sandp23/formal.tex
update.

This section describes the formal model of \systemname, called
\lang, making precise its syntax, semantics, and type system. It also
develops \lang's meta-theory, including type soundness, the non-exposure, and the non-crashing
theorem.

@@

There are two notions of type in \lang.  Types $\tau$ classify
word-sized values including integers and pointers, while types
$\omega$ classify multi-word values such as arrays, null-terminated
arrays, functions, and single-word-size values.
%
Pointer types ($\tptr{\omega}{m}$) include a mode annotation ($m$)
which is either checked (\cmode) or unchecked (\umode) and a type
($\omega$) denoting valid values that can be pointed to. Array types include both the type of
elements ($\tau$) and a bound ($\bvar$) comprised of an upper and
lower bound on the size of the array ($(b_l,b_h)$). Bounds $b$ are
limited to integer literals $n$ and expressions $x + n$.
Whether an array pointer is null terminated or not is determined by annotation
$\kappa$, which is $nt$ for null-terminated arrays, and $\cdot$
otherwise (we elide the $\cdot$ when writing the type).
\systemname function types ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$)
reflect its dependent function declrations,
where $\overline{x}$ represents
a list of \tint type variables in a dependent function header,
which bind all bounds appearing in $\overline{\tau}$ and $\tau$.
We have a well-formed requirement for a function type,
such that all variables in $\overline{\tau}$ and $\tau$ are bounded by $\overline{x}$.
Here is the
corresponding \systemname syntax for these types:
\[\hspace*{-0.5em}
\begin{array}{l}
\begin{array}{rcl}
$\code{array_tptr<}$\tau$\code{> : count(}$n$\code{)}$
&\Leftrightarrow& \tarrayptr{0}{n}{\tau}{\tmode}
\\[0.2em]
$\code{nt_array_ptr<}$\tau$\code{> : count(}$n$\code{)}$
&\Leftrightarrow& \tntarrayptr{0}{n}{\tau}{\cmode}
\end{array}
\\[0.2em]
$\code{tptr<(int)(nt_array_tptr<}$\tau$\code{> : count(}$n$\code{),}$\\
\qquad\qquad$\code{nt_array_tptr<}$\tau$\code{>)>: count(}$n$\code{))>}$
\\[0.2em]
\Leftrightarrow\;\; $\tptr{(\tfun{n}{ \tntarrayptr{0}{n}{\tau}{\tmode} \times \tntarrayptr{0}{n}{\tau}{\tmode}}{\tint})}{\tmode}$
\end{array}
\]
As a convention we write $\tptr{\tarrayb{b}{\tau}}{\cmode}$ to mean
$\tptr{\tarray{0}{b}{\tau}}{\cmode}$, so the above examples could
be rewritten $\tptr{\tarrayb{n}{\tau}}{\cmode}$ and
$\tptr{\tntarrayb{n}{\tau}}{\cmode}$, respectively.

@@

Integer literals $n$ are annotated with a type $\tau$ which can be either
$\tint$, or $\tptr{\omega}{m}$ in the case $n$ is being used as 
a heap address (this is useful for the semantics);
$\evalue{0}{\tptr{\omega}{m}}$ (for any $m$ and $\omega$) represents the $\enull$ pointer, as usual. 
The $\texttt{strlen}$ expression operates on variables $x$
rather than arbitrary expressions to simplify managing
bounds information in the type system; the more general case can be
encoded with a \code{let}. We use a less verbose syntax for dynamic bounds
casts; e.g., the following %

@@

Compared to the former \checkedc model \cite{li22checkedc},
there are four differences.
First, the \systemname type annotations have well-formed restrictions in \Cref{fig:wftypes}, 
for maintaining non-exposure.
Mainly, in a nested pointer $\tptr{(... \tptr{\tau}{\xi'} ...)}{\xi}$, if the outside mode $\xi$ is 
$\tmode$ or $\umode$, the inside mode $\xi'$ cannot be $\cmode$.
It is worth noting that pointer modes are a three point partial order ($\le$),
where $\tmode$ is the infimum, and
$\xi\wedge m$ is a special meet operation that projects pointer modes onto context modes,
such that a pointer mode $\tmode$ is projected as $\umode$.
Second, $\emalloc{\xi}{\omega}$ includes a mode flag
$\xi$ for allocating different pointers in different heap mode regions.
For simplicity, we disallow $\omega$ to be a function type ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$),
Third, the first expression $e$ in a function call ($\ecall{e}{\overline{e}}$) represents a function pointer.
Fourth, we update $\eunchecked{\overline{x}}{e}$ 
to specify $\overline{x}$ restricting all free variables appearing in $e$.
Additionally, we create $\echeckedtext$ blocks indicating the context mode change from $\umode$ to $\cmode$.
Both the free variables in the $\euncheckedtext$ and $\echeckedtext$ blocks
and the return types cannot be $\cmode$ mode for maintaining non-exposure,
i.e., no checked pointer can be communicated between $\euncheckedtext$ and $\echeckedtext$ blocks.

\lang aims to be simple enough to work with, but powerful enough to
encode realistic \systemname idioms. For example, mutable local
variables can be encoded as immutable locals that point to the heap;
the use of \code{&} can be simulated with \code{malloc};
and loops can be encoded as recursive function calls. \code{struct}s are
not in Fig.~\ref{fig:checkc-syn} for space reasons, but they are
actually in our model, and developed in
\iftr
Appendix~\ref{appx:struct}.
\else
the supplemental report~\cite{checkedc-tech-report}.
\fi
C-style \code{union}s have no safe typing
in \checkedc, so we omit them.
Additional syntactic descrition is listed in the \checkedc formalism paper \cite{li22checkedc}.

@@

\begin{figure}
{\small
$\hspace*{-1.2em}
    \begin{array}{l}
    \begin{array}{lll}
\mu & ::= & \evalue{n}{\tau}\\
e & ::= & \ldots \mid \ret{x}{\mu}{e}\\
r & ::= & e \mid \enull \mid \ebounds\\
E &::=& \Box \mid \ebinop{E}{e} \mid \ebinop{\evalue{n}{\tau}}{E}\mid \ecast{\tau}{E} \mid \edyncast{\tau}{E} \mid\estar{E}\mid\eassign{E}{e}\\[0.2em]
&&\mid\eassign{\evalue{n}{\tau}}{E}\mid \elet{x}{E}{e}\mid \eif{E}{e}{e}\\[0.2em]
&&\mid \ecall{E}{\overline{e}} \mid \ecall{\evalue{n}{\tau}}{\overline{E}} \mid 
\eunchecked{\overline{x}}{E}
\mid \echecked{\overline{x}}{E}

@@

The operational semantics for \lang is defined as a small-step
transition relation with the judgment $ (\varphi,\heap,e)
\longrightarrow_m (\varphi',\heap',r)$. Here, $\varphi$ is a
\emph{stack} mapping from variables to values $\evalue{n}{\tau}$ and
$\heap$ is a \emph{heap} that is partitioned into two parts ($\cmode$ and $\umode$ regions), each of which
maps addresses (integer literals) to values $\evalue{n}{\tau}$.
If a pointer has mode $\cmode$, it lives in the $\cmode$ region; otherwise, it lives in the $\umode$ region.
We wrote $\heap(m,n)$ to retrieve the $n$-location heap value in the $m$ region,
and $\heap(m)[n\mapsto \mu]$ to update location $n$ with the value $\mu$.
It is worth noting that \systemname is not a fat-pointer system;
thus, in every heap update, the value type annotation remains the same through program executions.
Additionally, for both stack and heap, 
we ensure $\fv(\tau)=\emptyset$ for all the value type annotations $\tau$.

While heap bindings can change, stack bindings are immutable---once
variable $x$ is bound to $\evalue{n}{\tau}$ in $\varphi$, that binding will not
be updated; we can model mutable stack variables as pointers into the
mutable heap.
As mentioned, value $\evalue{0}{\tau}$
represents a $\enull$ pointer when $\tau$ is a pointer type;
correspondingly, $\heap(m,0)$ should always be undefined.
The relation steps to a \emph{result} $r$,
which is either an expression or a $\enull$ or $\ebounds$ failure,
representing a null-pointer dereference or out-of-bounds access,
respectively. Such failures are a \emph{good} outcome; stuck states
(non-value expressions that cannot transition to a result $r$)
characterize undefined behavior.
%
The context mode $m$ indicates whether the
stepped redex within $e$ was in a checked ($\cmode$) or
unchecked ($\umode$) region.

The rules for the main operational semantics
judgment---\emph{evaluation}---are given at the bottom of
Fig.~\ref{fig:c-context}.
The first rule takes an expression $e$, decomposes
it into an \emph{evaluation context} $E$ and a sub-expression $e'$
(such that replacing the hole $\Box$ in $E$ with $e'$ would yield
$e$), and then evaluates $e'$ according to the \emph{computation}
  relation $(\varphi,\heap,e') \longrightarrow (\varphi,\heap,e'')$,
whose rules are given in Fig.~\ref{fig:semantics}, discussed
shortly. 
The second rule describes the exception handling 
for possible crashing behaviors in unchecked region.
A program in $\umode$ mode can non-deterministically crash
and the \systemname sandbox mechanism recorvers
the program to a safe point ($\evalue{0}{\tau}$)
and continues with the existing program state.
There are other rules for describing the halts of evaluation to $\enull$ and $\ebounds$ states in \Cref{app:main}.

%\footnote{This approach is that of the PLT Redex model of \lang; the Coq
%development uses a slightly simpler syntax to achieve the same
%effect.}
% \review{the special case raises questions, e.g. why is this syntax-driven and
%   not type-driven? }
% \liyi{This describes the semantic transition rules. We are using context evaluation framework to define the transition rules as the $E$ definition in Fig.3. like $\frac{x \Rightarrow y}{x+z \Rightarrow y + z}$, I don't know how type-driven can help us define translation rules.  }
% \mwh{Don't follow the above. I don't see this ``context transition
%   rule'' anywhere, and I'm not sure how it would fire, if we had it.}
% \liyi{The comment seems to confuse the meaning of the text about the if-then-else rules. Making the rule specific will help. }
The $\mode$ function
determines the context mode when evaluating $e'$ based on the context $E$.
For any program execution, the top context mode always starts with $\cmode$ ($\mode(E)=\mode'(E,\cmode)$).
The result context mode is depeneded on where $\Box$ locates.
If it occurs within $(\eunchecked{\overline{x}}{E'})$ inside $E$ without
a surrounding $\echeckedtext$ block, then the mode is
$\umode$; if the exact opposite happens (within $(\echecked{\overline{x}}{E'})$ without surrounding $\euncheckedtext$ blocks),
the mode is $\cmode$; the other cases ($\mode'(\alpha({E}),m)=\mode'(E,m)$) are recursively defined by the above base cases ($\alpha$ represents any syntactic categories in \Cref{fig:c-context} other than $\Box$, $\euncheckedtext$, and $\echeckedtext$).
Evaluation contexts $E$ define a standard left-to-right evaluation order. (We explain the
$\ret{x}{\mu}{e}$ syntax shortly.)
% For a term $e$, $C$ is a context if any
% only if it contains a $\Box$ hole term in one of the subterm position
% in $e$. $\Box$ indicates the break point where we split $e$ into $C$
% and a subterm $e'$ such that $C[e'] = e$.  For a function call
% $\ecall{f}{\overline{e}}$, a valid context ($\overline{C}$) is defined
% as any one of the elements in the list $\overline{e}$ being a
% context. We show a rule of the upper level \checkedc semantics. The
% rule splits a term $e$ into a context $C$ and redex $e'$.  If subterm
% $e'$ is transitioned to $e_a$, $C[e']$ is transitioned to
% $C[e_a]$.

@@

    \inferrule[S-AssignArrC]{\heap(\cmode,n)=\evalue{n_a}{\tau_a}\\ 0 \in [n_l,n_h) }
      {(\varphi,\heap,\eassign{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{\cmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,(\heap,\cmode)[n \mapsto \evalue{n_1}{\tau_a}],\evalue{n_1}{\tau})}

@@

    \inferrule[S-AssignArrT]{\heap(\umode,n)=\evalue{n_a}{\tau_a}\\ 0 \in [n_l,n_h)
               \\  \emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n_1}{\tau} }
      {(\varphi,\heap,\eassign{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{\tmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,(\heap,\umode)[n \mapsto \evalue{n_1}{\tau_a}],\evalue{n_1}{\tau})}

@@

    \inferrule[S-RetEnd]{}{(\varphi,\heap,\ret{x}{\mu}{\mu'}) \longrightarrow (\varphi,\heap,\mu')}

@@

    \inferrule[S-RetCon]{\varphi(x)=\mu\\ (\varphi[x\mapsto \mu'],\heap,e) \longrightarrow (\varphi',\heap',e')}{(\varphi,\heap,\ret{x}{\mu'}{e}) \longrightarrow (\varphi'[x\mapsto \mu],\heap',\ret{x}{\varphi'(x)}{e'})}

    \inferrule[S-Unchecked]{}{(\varphi,\heap,\eunchecked{\overline{x}}{\mu}) \longrightarrow (\varphi,\heap,\mu)}

@@

    \inferrule[S-Checked]{}{(\varphi,\heap,\echecked{\overline{x}}{\mu}) \longrightarrow (\varphi,\heap,\mu)}

@@

Fig.~\ref{fig:semantics} shows selected rules for the computation
relation; we explain them with the help of the example in \Cref{fig:checkedc-example-1} and \Cref{fig:checkedc-example-2}.

@@

\myparagraph{Checked and Tainted Pointer Operations}
%
The rules for pointer related operations---\textsc{S-DefC},
\textsc{S-DefT}, \textsc{S-AssignArrC}, \textsc{S-AssignArrT},
\textsc{S-DefNull}, and \textsc{S-Cast}.
The first five are deference and assignment operations---illustrate how the semantics checks bounds.
Rule \textsc{S-DefNull} transitions attempted null-pointer
dereferences to $\enull$, whereas \textsc{S-DefC} dereferences a \cmode-mode
non-null (single) pointer.
When $\enull$ is returned by the
computation relation, the evaluation relation halts the entire
evaluation with $\enull$ (using a rule not shown in Fig.~\ref{fig:c-context}); it
does likewise when $\ebounds$ is returned (see \Cref{sec:rem-semantics}).
\textsc{S-AssignArrC} assigns to an array as long as 0 (the point of
dereference) is within the bounds designated by the pointer's annotation
and strictly less than the upper bound.

\textsc{S-DefT} and \textsc{S-AssignArrT} are similar rules to \textsc{S-DefC} and \textsc{S-AssignArrC} for tainted pointers.
Any dynamic heap use of a tainted pointer requires a verification.
Performing such a verification equates to performing a type check for a pointer constant in \Cref{fig:const-type}.
We explain this shortly in \Cref{sec:type-system}.
For now, the verification step, e.g. $\umode;\emptyset;\heap ; \emptyset \vdash\evalue{n_a}{\tau}$ in \textsc{S-DefC},
means we verify that the value $n_a$ has type $\tau$ in the $\umode$ region heap.

@@

\myparagraph{Unchecked and Checked Blocks}
%
Semantically, $\euncheckedtext$ and $\echeckedtext$ blocks 
(\textsc{S-Unchecked} and \textsc{S-Checked}) act as a classical C block structures.
The interesting part is their type rules in \Cref{sec:type-system}.

\myparagraph{Binding and Function Calls}
%
The semantics handles variable scopes using the special $\erettext$
form. \textsc{S-Let} evaluates to a configuration whose stack
is $\varphi$ extended with a binding for $x$, and whose expression is
$\ret{x}{\evalue{n}{\tau}}{e})$ which keeps $\varphi$ unchanged, but
remember $x$ with the new value $\evalue{n}{\tau}$ in the scope $e$.
Every time when Evaluation proceeds on $e$ (Rule \textsc{S-RetCon}),
we install the current stack value $\evalue{n}{\tau}$ for $x$ in $\varphi$,
after one-step evaluation is completed, we restore the result $\varphi'(x)$ 
in the result $\ret{x}{\varphi'(x)}{e'})$.
This procedure continues until it becomes a literal
$n\!:\!\tau$, in which case \textsc{S-RetEnd} remove the $\kw{ret}$ frame and evaluates to
$n\!:\!\tau$.

Function calls are handled by \textsc{S-FunC} and \textsc{S-FunT},
for \cmode and \tmode mode function pointers, respectively. 
A call to function $f$ causes $f$'s
definition to be retrieved from $\Xi$,
% LEO: This is the first occurence of \Xi. It is explained,
% but reads a bit awkward as it's unclear this is its "introduction"
which maps function names to
function data $\tau\;(\evalue{\overline{x}}{\overline{\tau}})\;(\xi,e)$, where
$\tau$ is the return type, $(\evalue{\overline{x}}{\overline{\tau}})$
is the parameter list of variables and their types, 
$\xi$ determines the mode of the function, and $e$ is the
function body. 
Similar to \heap, the global function store $\Xi$ is also partitioned into
two parts ($\cmode$ and $\umode$ regions), each of which
maps addresses (integer literals) to the function data described above.

>>>>

This section describes the formal model of \systemname, named
\lang, making precise its syntax, semantics, and type system. It also
develops \lang's meta-theory, including the type soundness, non-exposure, and non-crashing
theorems.

@@

There are two type notions in \lang.  Types $\tau$ classify
word-sized values including integers and pointers, while types
$\omega$ classify multi-word values such as arrays, null-terminated
arrays, functions, and single-word-size values.
%
Pointer types ($\tptr{\omega}{\xi}$) include a pointer mode annotation 
($\xi$, the difference between context and pointer modes is introduced shortly below)
that is either checked (\cmode), tainted (\tmode), or unchecked (\umode),
and a type ($\omega$) denoting valid values that can be pointed to.
Array types include both the type of
elements ($\tau$) and a bound ($\bvar$) comprised of an upper and
lower bound on the size of the array ($(b_l,b_h)$). Bounds $b$ are
limited to integer literals $n$ and expressions $x + n$.
Whether an array pointer is null terminated or not is determined by annotation
$\kappa$, which is $nt$ for null-terminated arrays, and $\cdot$
otherwise (we elide $\cdot$ when writing types).
\lang function types ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$)
reflect its dependent function declarations,
where $\overline{x}$ represents
a list of \tint type variables in a dependent function header
that bind bound variables appearing in $\overline{\tau}$ and $\tau$.
We have a well-formed requirement for a function type,
such that all variables in $\overline{\tau}$ and $\tau$ are bounded by $\overline{x}$.
Here is the
corresponding \systemname syntax for these types:
\[\hspace*{-0.5em}
\begin{array}{l}
\begin{array}{rcl}
$\code{array_tptr<}$\tau$\code{> : count(}$n$\code{)}$
&\Leftrightarrow& \tarrayptr{0}{n}{\tau}{\tmode}
\\[0.2em]
$\code{nt_array_ptr<}$\tau$\code{> : count(}$n$\code{)}$
&\Leftrightarrow& \tntarrayptr{0}{n}{\tau}{\cmode}
\end{array}
\\[0.2em]
$\code{tptr<(int)(nt_array_tptr<}$\tau$\code{> : count(}$n$\code{),}$\\
\qquad\qquad$\code{nt_array_tptr<}$\tau$\code{>)>: count(}$n$\code{))>}$
\\[0.2em]
\Leftrightarrow\;\; $\tptr{(\tfun{n}{ \tntarrayptr{0}{n}{\tau}{\tmode} \times \tntarrayptr{0}{n}{\tau}{\tmode}}{\tint})}{\tmode}$
\end{array}
\]
As a convention we write $\tptr{\tarrayb{b}{\tau}}{\cmode}$ to mean
$\tptr{\tarray{0}{b}{\tau}}{\cmode}$, so the above examples could
be rewritten $\tptr{\tarrayb{n}{\tau}}{\cmode}$ and
$\tptr{\tntarrayb{n}{\tau}}{\cmode}$, respectively.

@@

Integer literals $n$ are annotated with a type $\tau$ which can be either
$\tint$, or $\tptr{\omega}{\xi}$ in the case $n$ is being used as 
a heap address (this is useful for the semantics);
$\evalue{0}{\tptr{\omega}{\xi}}$ (for any $\xi$ and $\omega$) represents the $\enull$ pointer, as usual. 
The $\texttt{strlen}$ expression operates on variables $x$
rather than arbitrary expressions to simplify managing
bounds information in the type system; the more general case can be
encoded with a \code{let}. We use a less verbose syntax for dynamic bounds
casts; e.g., the following %

@@

Compared to the former \checkedc model \cite{li22checkedc},
there are four differences.
First, the \systemname type annotations have well-formed restrictions in \Cref{fig:wftypes}, 
for maintaining non-exposure.
Mainly, in a nested pointer $\tptr{(... \tptr{\tau}{\xi'} ...)}{\xi}$, $\xi'\le \xi$.
It is worth noting that pointer modes are a three point partial order ($\le$),
where $\tmode$ is the infimum, and
$\xi\wedge m$ is a special meet operation that projects pointer modes onto context modes,
such that $\tmode$ is projected as $\umode$.
Second, $\emalloc{\xi}{\omega}$ includes a mode flag
$\xi$ for allocating different pointers in different heap mode regions.
We disallow $\omega$ to be a function type ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$).
Third, the first expression $e$ in a function call ($\ecall{e}{\overline{e}}$) represents a function pointer.
Fourth, $\echeckedtext$ blocks are added to the system, 
which permits the nested switching between $\echeckedtext$ (represented by context mode $\cmode$)
and $\euncheckedtext$ (represented by context mode $\umode$) code regions.
One example usage of the nested switching is the checked function callbacks inside 
an unsafe region in \Cref{fig:checkedc-example-2} and \ref{fig:checkedc-example-3}.
To guarantee the non-exposure safety,
 we extend the $\echeckedtext$ and $\euncheckedtext$ block syntax to be 
$\echecked{\overline{x}}{e}$ and $\eunchecked{\overline{x}}{e}$:
$\overline{x}$ restricts all free variables appearing in $e$, and they cannot be checked pointers.

\lang aims to be simple enough to work with, but powerful enough to
encode realistic \systemname idioms. For example, mutable local
variables can be encoded as immutable locals that point to the heap;
the use of \code{&} can be simulated with \code{malloc};
and loops can be encoded as recursive function calls. \code{struct}s are
not in Fig.~\ref{fig:checkc-syn} for space reasons, but they are
actually in our model, and developed in
\iftr
Appendix~\ref{appx:struct}.
\else
the supplemental report~\cite{checkedc-tech-report}.
\fi
C-style \code{union}s have no safe typing
in \checkedc, so we omit them.

@@

\begin{figure}
{\small
$\hspace*{-1.2em}
    \begin{array}{l}
    \begin{array}{lll}
e & ::= & \ldots \mid \ret{x}{\evalue{n}{\tau}}{e}\\
r & ::= & e \mid \enull \mid \ebounds\\
E &::=& \Box \mid \ebinop{E}{e} \mid \ebinop{\evalue{n}{\tau}}{E}\mid \ecast{\tau}{E} \mid \edyncast{\tau}{E} \mid\estar{E}\mid\eassign{E}{e}\\[0.2em]
&&\mid\eassign{\evalue{n}{\tau}}{E}\mid \elet{x}{E}{e}\mid \eif{E}{e}{e}\\[0.2em]
&&\mid \ecall{E}{\overline{e}} \mid \ecall{\evalue{n}{\tau}}{\overline{E}} \mid 
\eunchecked{\overline{x}}{E}
\mid \echecked{\overline{x}}{E}

@@

The operational semantics for \lang is defined as a small-step
transition relation with the judgment $ (\varphi,\heap,e)
\longrightarrow_m (\varphi',\heap',r)$. Here, $\varphi$ is a
\emph{stack} mapping from variables to values $\evalue{n}{\tau}$ and
$\heap$ is a \emph{heap} that is partitioned into two parts ($\cmode$ and $\umode$ regions), each of which
maps addresses (integer literals) to values $\evalue{n}{\tau}$.
A $\cmode$pointer is mapped to a heap location in the $\cmode$ region, 
while a $\tmode$ and $\umode$ pointer represents a $\umode$ region location.
We wrote $\heap(m,n)$ to retrieve the $n$-location heap value in the $m$ region,
and $\heapup{m}{n}{\evalue{n'}{\tau}}$ 
to update location $n$ with the value $\evalue{n'}{\tau}$ in the $m$ region.
It is worth noting that \systemname is not a fat-pointer system;
thus, in every heap update, the value type annotation remains the same through program executions.
Additionally, for both stack and heap, 
we ensure $\fv(\tau)=\emptyset$ for all the value type annotations $\tau$.

While heap bindings can change, stack bindings are immutable---once
variable $x$ is bound to $\evalue{n}{\tau}$ in $\varphi$, that binding will not
be updated; we can model mutable stack variables as pointers into the
mutable heap.
As mentioned, value $\evalue{0}{\tau}$
represents a $\enull$ pointer when $\tau$ is a pointer type;
correspondingly, $\heap(m,0)$ should always be undefined.
The relation steps to a \emph{result} $r$,
which is either an expression or a $\enull$ or $\ebounds$ failure,
representing a null-pointer dereference or out-of-bounds access,
respectively. Such failures are a \emph{good} outcome; stuck states
(non-value expressions that cannot transition to a result $r$)
characterize undefined behavior.
%
The context mode $m$ (in $\longrightarrow_{m}$) indicates whether the
stepped redex within $e$ was in a checked ($\cmode$) or
unchecked ($\umode$) region.

The rules for the main operational semantics
judgment---\emph{evaluation}---are given at the middle of
Fig.~\ref{fig:c-context}.
The first rule takes an expression $e$, decomposes
it into an \emph{evaluation context} $E$ and a sub-expression $e'$
(such that replacing the hole $\Box$ in $E$ with $e'$ would yield
$e$), and then evaluates $e'$ according to the \emph{computation}
  relation $(\varphi,\heap,e') \longrightarrow (\varphi,\heap,e'')$,
whose rules are given in Fig.~\ref{fig:semantics}, discussed
shortly. 
The second rule describes the exception handling 
for possible crashing behaviors in unchecked region.
A $\umode$ mode operation can non-deterministically crash
and the \systemname sandbox mechanism recovers
the program to a safe point ($\evalue{0}{\tau}$)
and continues with the existing program state.
Evaluation contexts $E$ define a standard left-to-right evaluation order. (We explain the
$\ret{x}{\mu}{e}$ syntax shortly.)
There are other rules for describing the halts of evaluation to $\enull$ and $\ebounds$ states in \Cref{app:main}.

%\footnote{This approach is that of the PLT Redex model of \lang; the Coq
%development uses a slightly simpler syntax to achieve the same
%effect.}
% \review{the special case raises questions, e.g. why is this syntax-driven and
%   not type-driven? }
% \liyi{This describes the semantic transition rules. We are using context evaluation framework to define the transition rules as the $E$ definition in Fig.3. like $\frac{x \Rightarrow y}{x+z \Rightarrow y + z}$, I don't know how type-driven can help us define translation rules.  }
% \mwh{Don't follow the above. I don't see this ``context transition
%   rule'' anywhere, and I'm not sure how it would fire, if we had it.}
% \liyi{The comment seems to confuse the meaning of the text about the if-then-else rules. Making the rule specific will help. }
The $\mode$ function  at the bottom of Fig.~\ref{fig:c-context}
describes the context mode determination in each evaluation step based on the context $E$.
For any program execution, the function starts the mode computation with $\cmode$ ($\mode(E)=\mode'(E,\cmode)$).
The result context mode depends on where $\Box$ locates.
If it occurs within $E$ in $(\eunchecked{\overline{x}}{E})$
that has no surrounding $\echeckedtext$ block,
the mode is $\umode$; otherwise, the mode is $\cmode$.
$\mode'(\alpha({E}),m)=\mode'(E,m)$ represent other construct cases 
that are not $\echeckedtext$ and $\euncheckedtext$; in such case, 
the function recursively traverses the sub-context to find the context mode.

% For a term $e$, $C$ is a context if any
% only if it contains a $\Box$ hole term in one of the subterm position
% in $e$. $\Box$ indicates the break point where we split $e$ into $C$
% and a subterm $e'$ such that $C[e'] = e$.  For a function call
% $\ecall{f}{\overline{e}}$, a valid context ($\overline{C}$) is defined
% as any one of the elements in the list $\overline{e}$ being a
% context. We show a rule of the upper level \checkedc semantics. The
% rule splits a term $e$ into a context $C$ and redex $e'$.  If subterm
% $e'$ is transitioned to $e_a$, $C[e']$ is transitioned to
% $C[e_a]$.

@@

    \inferrule[S-AssignArrC]{\heap(\cmode,n)=\evalue{n_a}{\tau_a}\\ 0 \in [n_l,n_h) }
      {(\varphi,\heap,\eassign{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{\cmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,\heapup{\cmode}{n}{\evalue{n_1}{\tau_a}},\evalue{n_1}{\tau})}

@@

    \inferrule[S-AssignArrT]{\heap(\umode,n)=\evalue{n_a}{\tau_a}\\ 0 \in [n_l,n_h)
               \\  \emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n_1}{\tau} }
      {(\varphi,\heap,\eassign{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{\tmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,\heapup{\umode}{n}{\evalue{n_1}{\tau_a}},\evalue{n_1}{\tau})}

@@

    \inferrule[S-RetEnd]{}{(\varphi,\heap,\ret{x}{\evalue{n}{\tau}}{\evalue{n'}{\tau'}}) \longrightarrow (\varphi,\heap,\evalue{n'}{\tau'})}

@@

    \inferrule[S-RetCon]{ (\varphi[x\mapsto \evalue{n}{\tau}],\heap,e) \longrightarrow (\varphi',\heap',e')}{(\varphi,\heap,\ret{x}{\evalue{n}{\tau}}{e}) \longrightarrow (\varphi'[x\mapsto \varphi(x)],\heap',\ret{x}{\varphi'(x)}{e'})}

    \inferrule[S-Unchecked]{}{(\varphi,\heap,\eunchecked{\overline{x}}{\evalue{n}{\tau}}) \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}

@@

    \inferrule[S-Checked]{}{(\varphi,\heap,\echecked{\overline{x}}{\evalue{n}{\tau}}) \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}

@@

Fig.~\ref{fig:semantics} shows selected rules for the computation relation.

@@

\myparagraph{Checked and Tainted Pointer Operations}
%
The rules for pointer related operations---\textsc{S-DefC},
\textsc{S-DefT}, \textsc{S-AssignArrC}, \textsc{S-AssignArrT},
\textsc{S-DefNull}, and \textsc{S-Cast}.
The first five define deference and assignment operations---illustrate how the semantics checks bounds.
Rule \textsc{S-DefNull} transitions attempted null-pointer
dereferences to $\enull$, whereas \textsc{S-DefC} dereferences a $\cmode$-mode
non-null (single) pointer.
When $\enull$ is returned by the
computation relation, the evaluation relation halts the entire
evaluation with $\enull$ (using a rule not shown in Fig.~\ref{fig:c-context}); it
does likewise when $\ebounds$ is returned (see \Cref{sec:rem-semantics}).
\textsc{S-AssignArrC} assigns to an array as long as 0 (the point of
dereference) is within the bounds designated by the pointer's annotation
and strictly less than the upper bound.

\textsc{S-DefT} and \textsc{S-AssignArrT} are similar rules to \textsc{S-DefC} and \textsc{S-AssignArrC} for tainted pointers.
Any dynamic heap use of a tainted pointer requires a verification.
Performing such a verification equates to performing a type check for a pointer constant in \Cref{fig:const-type}.
We explain this shortly in \Cref{sec:type-system}.
For now, the verification step, e.g. $\emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n_a}{\tau}$ in \textsc{S-DefC},
means we verify that the value $n_a$ is well-defined in $\heap(m,n_a)$ and has type $\tau$, if $\tau$ is a pointer.

@@

\myparagraph{Unchecked and Checked Blocks}
%
Semantically, $\euncheckedtext$ and $\echeckedtext$ blocks act as classical C blocks
as rules \textsc{S-Unchecked} and \textsc{S-Checked} in \Cref{fig:semantics}.

\myparagraph{Binding and Function Calls}
%
The semantics handles variable scopes using the special $\erettext$
form. \textsc{S-Let} evaluates to a configuration whose expression is
$\ret{x}{\evalue{n}{\tau}}{e})$. We keep $\varphi$ unchanged
and remember $x$ and its the new value $\evalue{n}{\tau}$
in $e$'s scope that is defined by the $\erettext$ operation.
Every time when evaluation proceeds on $e$ (rule \textsc{S-RetCon}),
we install the stack value $\evalue{n}{\tau}$ for $x$ in $\varphi$ for the current scope;
after one-step evaluation is completed, 
we store $x$'s change in the result $\erettext$ operation $\ret{x}{\varphi'(x)}{e'})$,
and restore $x$'s outer score value $\varphi(x)$ in $\varphi'$. 
This procedure continues until $e'$ becomes a literal
$n\!:\!\tau$, in which case \textsc{S-RetEnd} removes the $\kw{ret}$ frame and returns
the literal.

Function calls are handled by \textsc{S-FunC} and \textsc{S-FunT},
for $\cmode$ and $\tmode$ mode function pointers, respectively. 
A call to a function pointer $n$ retrieves
 the function definition in $n$'s location in the global function store $\Xi$,
% LEO: This is the first occurence of \Xi. It is explained,
% but reads a bit awkward as it's unclear this is its "introduction"
which maps function pointers to
function data $\tau\;(\evalue{\overline{x}}{\overline{\tau}})\;(\xi,e)$, where
$\tau$ is the return type, $(\evalue{\overline{x}}{\overline{\tau}})$
is the parameter list of variables and their types, 
$\xi$ determines the mode of the function, and $e$ is the
function body. 
Similar to \heap, the global function store $\Xi$ is also partitioned into
two parts ($\cmode$ and $\umode$ regions), each of which
maps addresses (integer literals) to the function data described above.
